---
title: "superstore_data_analysis"
author: "Ziting Liang"
date: '2022-04-25'
output: pdf_document
---

## Import packages
```{r,message=FALSE}
library(dplyr)
library(car)
library(ggplot2)
```

## Read data
```{r}
Data<-read.csv('./Sample - Superstore.csv',header=T)
summary(Data)
#apply(is.na(Data),2,sum)#No NA

#check for some key variables
#unique(Data$Variable.name)

#select useful variables
data.clean<-Data%>%select(Order.Date,Ship.Mode,Customer.ID,Segment,State,Region,Category,
                          Sub.Category,Sales,Quantity,Discount,Profit)
summary(data.clean)
```
## Variable recoding
```{r}
#recode 4 new variables for Ship.Mode
data.clean$Ship.Mode[data.clean$Ship.Mode=='Second Class']<-1
data.clean$Ship.Mode[data.clean$Ship.Mode=='Standard Class']<-2
data.clean$Ship.Mode[data.clean$Ship.Mode=='First Class']<-3
data.clean$Ship.Mode[data.clean$Ship.Mode=='Same Day']<-4

#recode 3 new variables for Segment
data.clean$Segment[data.clean$Segment=='Consumer']<-1
data.clean$Segment[data.clean$Segment=='Corporate']<-2
data.clean$Segment[data.clean$Segment=='Home Office']<-3

#recode 4 new variables for Region
data.clean$Region[data.clean$Region=='South']<-1
data.clean$Region[data.clean$Region=='West']<-2
data.clean$Region[data.clean$Region=='Central']<-3
data.clean$Region[data.clean$Region=='East']<-4

#recode 3 new variables for Category
data.clean$Category[data.clean$Category=='Furniture']<-1
data.clean$Category[data.clean$Category=='Office Supplies']<-2
data.clean$Category[data.clean$Category=='Technology']<-3

data.clean$Sales<-round(data.clean$Sales, digits=2) 
data.clean$Profit<-round(data.clean$Profit, digits=2)

data.clean$Order.Date<-as.Date(data.clean$Order.Date,format='%m/%d/%Y')
```


## Descriptive statistics
```{r}
# pie chart of region
data_region<-data.clean %>% select(Region)
slices <- table(data.clean$Region)
count<-slices/sum(slices)*100
lbls <- c("South", "West", "Central", 'East')
pie(slices, col=c('pink','lightgreen','lightblue','lightyellow'),
    labels = paste0(lbls,',',round(count,2),'%'), main="Pie Chart of Region")

# pie chart of Ship.Mode
data_region<-data.clean %>% select(Ship.Mode)
slices <- table(data.clean$Ship.Mode)
count<-slices/sum(slices)*100
lbls <- c("Second Class", "Standard Class", "First Class", 'Same Day')
pie(slices, col=c('pink','lightgreen','lightblue','lightyellow'),
    labels = paste0(lbls,',',round(count,2),'%'), main="Pie Chart of Ship.Mode")

#A comparison of Sales and Profit in 2014-2017
chrono_data<-data.clean%>%group_by(Order.Date)%>%summarize(Sales=sum(Sales),Profit=sum(Profit))
matplot(chrono_data$Order.Date, cbind(chrono_data$Sales,chrono_data$Profit),
        type="l",col=c("orange","navyblue"),lty=c(1,1),xlab='Time',ylab='Sales/Profit',
        main='Sales and Profit in 2014-2017',axes=F,
        xlim=c(as.Date('2014-01-01'),as.Date('2018-01-01')))
Axis(side=2)
legend(x='topright',legend=c('Sales','Profit'),lty=1,col=c('orange','navyblue'),bty='n')
axis.Date(1,at=seq(from=as.Date('2014-01-01'),to=as.Date('2018-01-01'),by='3 months'),
          format='%y-%m',las=3)
```

## Convert to superstore.reg: preparation for regression
```{r}
#data.clean$Order.Date<-as.Date(data.clean$Order.Date,format='%m%d%Y')

#turn category variables to numeric variables
data.clean$Ship.Mode<-as.numeric(data.clean$Ship.Mode)
data.clean$Segment<-as.numeric(data.clean$Segment)
data.clean$Region<-as.numeric(data.clean$Region)
data.clean$Category<-as.numeric(data.clean$Category)

superstore.reg<-data.clean%>%select(Ship.Mode,Segment,Region,Category,
                                    Sales,Quantity,Discount,Profit)

#create 4 new variables for Ship.Mode for regression in case of multicollinearity, 
#and then remove Ship.Mode.
superstore.reg<-superstore.reg%>%
  mutate(SM_second_class=as.numeric(Ship.Mode==1),
         SM_standard_class=as.numeric(Ship.Mode==2),
         SM_first_class=as.numeric(Ship.Mode==3))%>%
  select(-Ship.Mode)

#create 3 new variables for Segment for regression in case of multicollinearity, 
#and then remove Segment.
superstore.reg<-superstore.reg%>%mutate(Seg_cons=as.numeric(Segment==1),
                                        Seg_corp=as.numeric(Segment==2))%>%select(-Segment)

#create 4 new variables for Region for regression in case of multicollinearity, 
#and then remove Region.
superstore.reg<-superstore.reg%>%mutate(Reg_south=as.numeric(Region==1),
                                        Reg_west=as.numeric(Region==2),
                                        Reg_central=as.numeric(Region==3))%>%select(-Region)

#create 3 new variables for Category for regression in case of multicollinearity,
#and then remove Category.
superstore.reg<-superstore.reg%>%mutate(Cat_furniture=as.numeric(Category==1),
                                        Cat_office=as.numeric(Category==2))%>%select(-Category)
```

## Regression before scale (including Correlation Analysis)
```{r}
###Regression

#Do a normality test before the regression
hist(superstore.reg$Sales, col='pink', main='Sales', breaks=50)
hist(superstore.reg$Profit, col='pink', main='Profit', breaks=50)
hist(superstore.reg$Discount, col='pink', main='Discount',breaks=50)

qqnorm(superstore.reg$Sales, main='Sales')
qqline(superstore.reg$Sales)

qqnorm(superstore.reg$Profit, main='Profit')
qqline(superstore.reg$Profit)

qqnorm(superstore.reg$Discount, main='Discount')
qqline(superstore.reg$Discount)

#Do regression and see what's going on
lm_fit<-lm(Sales~.,superstore.reg)
summary(lm_fit)#not fit well
reduced<-lm(Sales~1,superstore.reg)
full<-lm(Sales~.,superstore.reg)
step(reduced,scope=c(lower=reduced,upper=full),direction='forward',trace=F)
step(full,scope=c(lower=reduced,upper=full),direction='backward',trace=F)

#after doing regression, we want to do a correlation analysis to see if all variables are uncorrelated:
corrplot::corrplot(cor(
  superstore.reg%>%select(c(Quantity,Discount,Profit,Reg_central,Cat_furniture,Cat_office))),
  method='number')
#we want to see what will the model perform if we delete one of cat_furniture or cat_office
summary(lm(Sales~Quantity+Discount+Profit+Reg_central+Cat_furniture,data=superstore.reg))
summary(lm(Sales~Quantity+Discount+Profit+Reg_central+Cat_office,data=superstore.reg))
```

## Regression after scale or preprocessing (including Correlation Analysis)
```{r}
#normalization for profit
superstore.reg.scale<-superstore.reg
superstore.reg.scale$Profit<-scale(superstore.reg.scale$Profit)
#take logarithm for sales
superstore.reg.scale$Sales<-log(superstore.reg.scale$Sales+1)

#Do a normality test before the regression
hist(superstore.reg.scale$Sales, col='pink', main='Sales', breaks = 50)
hist(superstore.reg.scale$Profit, col='pink', main='Profit',breaks = 50)
hist(superstore.reg.scale$Discount, col='pink', main='Discount',breaks = 50)

qqnorm(superstore.reg.scale$Sales, main='Sales')
qqline(superstore.reg.scale$Sales)

qqnorm(superstore.reg.scale$Profit, main='Profit')
qqline(superstore.reg.scale$Profit)

qqnorm(superstore.reg.scale$Discount, main='Discount')
qqline(superstore.reg.scale$Discount)

#Do regression and see what's going on
lm_fit<-lm(Sales~.,superstore.reg.scale)
summary(lm_fit)#not fit well
reduced<-lm(Sales~1,superstore.reg.scale)
full<-lm(Sales~.,superstore.reg.scale)
step(reduced,scope=c(lower=reduced,upper=full),direction='forward',trace=F)
step(full,scope=c(lower=reduced,upper=full),direction='backward',trace=F)

#after doing regression, we want to do a correlation analysis to see if all variables are uncorrelated:
corrplot::corrplot(cor(
  superstore.reg.scale%>%
    select(c(Quantity,Discount,Profit,Reg_central,Cat_furniture,Cat_office))),
  method='number')

#we want to see what will the model perform if we delete one of cat_furniture or cat_office
summary(lm(Sales~Quantity+Discount+Profit+Cat_furniture,data=superstore.reg.scale))
summary(lm(Sales~Quantity+Discount+Profit+Cat_office,data=superstore.reg.scale))

#Regression tree for scaled
library(rpart)
rt<-rpart(Sales~Quantity+Discount+Profit+Cat_office,superstore.reg.scale)
par(xpd = TRUE)
plot(rt, compress = TRUE)
text(rt, use.n = TRUE,cex=0.55)
```

## ANCOVA
```{r}
#transfer data
ancova_data<-data.clean %>%select(Sales,Profit,Discount)
idx<-1:9994
disc_No<-idx[ancova_data$Discount==0]
disc_Low<-idx[ancova_data$Discount>0&ancova_data$Discount<=0.2]
disc_High<-idx[ancova_data$Discount>0.2]
ancova_data$Discount[disc_No]<-'No'
ancova_data$Discount[disc_Low]<-'Low'
ancova_data$Discount[disc_High]<-'High'

rm(disc_No,disc_Low,disc_High,idx)
#EDA and summary
ancova_data %>%
  group_by(Discount) %>%  
  summarise(mean_sales = mean(Sales),
            median_sales = median(Sales),
            sd_sales = sd(Sales),
            mean_profit = mean(Profit),
            median_profit = median(Profit),
            sd_profit = sd(Profit))
#boxplot
boxplot(Profit ~ Discount,data = ancova_data,main = "Profit by Discount",
        xlab = "Discount",ylab = "Profit",col = "blue",border = "pink")
boxplot(Sales ~ Discount,data = ancova_data,main = "Sales by Discount",
        xlab = "Discount",ylab = "Sales",col = "blue",border = "pink")

#hypothesis testing:independence btw discount and sales, and equality of variance
summary(aov(Sales ~ Discount,data = ancova_data))
#The p-value is 0.136 that is greater than 0.05, so Discount and Sales are independent to each other.

#Levene’s Test
leveneTest(Profit~Discount, data = ancova_data)
#The p-value of the test is 2.2e-16, which indicates that the variances among the groups are not equal.

#Fit analysis of covariance model ANCOVA
ancova_model <- aov(Profit ~ Discount + Sales, data = ancova_data)
Anova(ancova_model, type="III")
```
From this result, we can conclude that no matter we control discount or sales still, the other variable is significant in this model, which indicates that they will significantly contribute to the model. From two significant p-value of Discount and Sales, we may say that: (1) different Discount will influence the relationship between Sales and Profit, and (2) sales does have relationship to profit.

```{r}
lm_ancova<-lm(Profit~Sales+Sales:Discount,data=ancova_data)
summary(lm_ancova)
```
From this linear model with interaction of Sales and Discount, we may say that different discount will contribute different relationship between Sales and Profit. Here, when discount is 0, the coefficient is 0.5684-0.2032=0.3652; when discount is low, the coefficient is 0.3726-0.2032=0.1694; when discount is high, the coefficient is -0.2032.

```{r}
tableone<-data.frame(ancova_data$Discount, data.clean$Category)
slices <- table(tableone$ancova_data.Discount)
count<-slices/sum(slices)*100
lbls <- c("High", "Low", "No")
pie(slices, col=c('pink','brown','purple'),labels = paste0(lbls,',',round(count,2),'%'), 
    main="Pie Chart of Discount")

tableone_high<-tableone %>% filter(ancova_data$Discount=='High')
slices <- table(tableone_high$data.clean.Category)
count<-slices/sum(slices)*100
lbls <- c("Furniture", "Office Supplies", "Technology")
pie(slices, col=c('pink','brown','purple'),labels = paste0(lbls,',',round(count,2),'%'), 
    main="Pie Chart of Categories with High Discount")


```




## RFM analysis
```{r}
#create RFM table
RFM.data<-data.clean[lubridate::year(data.clean$Order.Date)%in%c(2016,2017),]
RFM.table<-RFM.data%>%
  group_by(Customer.ID)%>%
  summarize(Recency=max(Order.Date),Frequency=n(),Monetary=sum(Sales))
RFM.table$Recency<-as.numeric((as.Date('2017-12-31')-RFM.table$Recency))
summary(RFM.table)

#Scored data
rfm_scored<-RFM.table
rfm_scored$r_score <- rep(0,773)
rfm_scored$r_score[RFM.table$Recency >= 163.0] <- 1
rfm_scored$r_score[RFM.table$Recency >=73.0 & RFM.table$Recency < 163.0] <- 2
rfm_scored$r_score[RFM.table$Recency >=30.0 & RFM.table$Recency < 73.0] <- 3
rfm_scored$r_score[RFM.table$Recency < 30.0] <- 4

rfm_scored$f_score <- rep(0,773)
rfm_scored$f_score[RFM.table$Frequency >= 10.000] <- 1
rfm_scored$f_score[RFM.table$Frequency >=7.000 & RFM.table$Frequency < 10.000] <- 2
rfm_scored$f_score[RFM.table$Frequency >=4.000 & RFM.table$Frequency < 7.000] <- 3
rfm_scored$f_score[RFM.table$Frequency < 4.000] <- 4

rfm_scored$m_score <- rep(0,773)
rfm_scored$m_score[RFM.table$Monetary >= 2216.88] <- 1
rfm_scored$m_score[RFM.table$Monetary >=1194.96 & RFM.table$Monetary < 2216.88] <- 2
rfm_scored$m_score[RFM.table$Monetary >=519.76 & RFM.table$Monetary < 1194.96] <- 3
rfm_scored$m_score[RFM.table$Monetary <519.76] <- 4

rfm_scored<-rfm_scored %>% 
  mutate(RFM_score=r_score*100+f_score*10+m_score) %>%
  select(Customer.ID,r_score,f_score,m_score,RFM_score)

#segments
rfm_scored$Segment <- "0"
rfm_scored$Segment[which(rfm_scored$RFM_score 
  %in% c(444,434,443, 344, 442, 244, 424, 441))] <-"Loyalists"
rfm_scored$Segment[which(rfm_scored$RFM_score 
  %in% c(332,333,342,343,334,412,413,414,431,432,441,421,422,423,424,433))]<- "Potential Loyalists"
rfm_scored$Segment[which(rfm_scored$RFM_score 
  %in% c(233,234, 241,311, 312, 313,314,321,322,323,324, 331,  341))] <- "Promising"
rfm_scored$Segment[which(rfm_scored$RFM_score 
  %in% c(124, 133, 134, 142, 143, 144, 214,224,234, 242, 243, 232 ))] <- "Hesitant"
rfm_scored$Segment[which(rfm_scored$RFM_score 
  %in% c(122, 123,131 ,132, 141, 212, 213, 221, 222, 223, 231 ))] <- "Need attention"
rfm_scored$Segment[which(rfm_scored$RFM_score 
  %in% c(111, 112, 113, 114, 121, 131, 211, 311, 411 ))] <-"Detractors"

#plot of segments
rfm_scored%>%
  group_by(Segment)%>%
  summarize(Count=n())%>%
  ggplot(aes(x = forcats::fct_reorder(Segment, Count),y=Count,fill = Segment)) + 
  geom_bar(stat='identity')+
  geom_text(aes(label=Count),nudge_y=-.5,color="white",size = 3.5,vjust=1.2)+
  theme(axis.text.x=element_text(angle=30,hjust=1))+
  labs(title = "Barplot for Segments of customers")



```




